---
title: "gapminder"
output: html_document
---

Download and load the gapminder data.

```{r}
url <- "http://www.stat.ubc.ca/~jenny/notOcto/STAT545A/examples/gapminder/data/gapminderDataFiveYear.txt"
download.file(url, "gapminderDataFiveYear.txt", method="curl")
df <- read.table("gapminderDataFiveYear.txt", head=T, sep="\t")
df <- read.delim("gapminderDataFiveYear.txt")
```

Thank god that was properly formatted.

Getting to know a `data.frame`

```{r}
str(df)
head(df)
tail(df)
summary(df)
names(df)
dim(df)
nrow(df)
length(df)
ncol(df)
```

* `data.frames` are actually a `list()` R Object.
* they can hold many different kinds of R Object, recall that all other data types in R are `atomic`
* list that acts like a matrix, each of these lists has to be the same length
* generic functions have actions that depend on the type of object handed to it
* `summary()` 
  * allows you to see if there are problems
  * looks 

Now we will load the `ggplot2` library.

```{r}
library("ggplot2")
qplot(data=df, lifeExp, log(gdpPercap), col=continent) + theme_bw()
```