---
title: "gapminder"
output: html_document

---

Download and load the gapminder data.

```{r}
setwd("~/Dropbox/courses/stat545A/gapminder/")
url <- "http://www.stat.ubc.ca/~jenny/notOcto/STAT545A/examples/gapminder/data/gapminderDataFiveYear.txt"
download.file(url, "gapminderDataFiveYear.txt", method="curl")
df <- read.table("gapminderDataFiveYear.txt", head=T, sep="\t", fill= TRUE )
df <- read.delim("gapminderDataFiveYear.txt")
```

Thank god that was properly formatted.

Getting to know a `data.frame`

```{r}
str(df)
head(df)
tail(df)
summary(df)
names(df)
dim(df)
nrow(df)
length(df)
ncol(df)
```

* `data.frames` are actually a `list()` R Object.
* they can hold many different kinds of R Object, recall that all other data types in R are `atomic`
* list that acts like a matrix, each of these lists has to be the same length
* generic functions have actions that depend on the type of object handed to it
* `summary()` 
  * allows you to see if there are problems
  * give you a quick idea of current data stories 

Now we will load the `ggplot2` library.

```{r}
library("ggplot2")
# nothing to print
ggplot(df, aes(x = gdpPercap, y = lifeExp)) + geom_point()

```



## Factors

* special type of structure categorial variable
* values stored as integers with `levels` that map to the chatacter strings and their relative order

```{r}
str(df$continent)
nlevels(df$continent)
levels(df$continent)
table(df$continent)
barplot(table(df$continent))
```